---
title: "Data Analysis Project"
author: "FateenAhmed"
date: "2024-02-25"
header-includes:
  - \usepackage{sectsty}
  - \sectionfont{\clearpage}
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width = 5, fig.height = 3)
```

## Reading and Understanding the Datasets

### Air Quality System (AQS) Dataset

The AQS dataset provides comprehensive data on air quality across the United States across 4 years.

```{r get-aqs-data}
library(readr)
aqs_data_2021 <- read_csv("data/aqs/annual_conc_by_monitor_2021.csv")
head(aqs_data_2021,5)
aqs_data_2020 <- read_csv("data/aqs/annual_conc_by_monitor_2020.csv")
head(aqs_data_2020,5)
aqs_data_2019 <- read_csv("data/aqs/annual_conc_by_monitor_2019.csv")
head(aqs_data_2019,5)
aqs_data_2018 <- read_csv("data/aqs/annual_conc_by_monitor_2018.csv")
head(aqs_data_2018,5)
```

```{r get-aqs-info}
# Listing columns for AQS datasets
cat("Columns for AQS 2021:\n", paste(names(aqs_data_2021), collapse = ", "), "\n\n")
cat("Columns for AQS 2020:\n", paste(names(aqs_data_2020), collapse = ", "), "\n\n")
cat("Columns for AQS 2019:\n", paste(names(aqs_data_2019), collapse = ", "), "\n\n")
cat("Columns for AQS 2018:\n", paste(names(aqs_data_2018), collapse = ", "), "\n\n")

```

### Behavioral Risk Factor Surveillance System (BRFSS) Dataset

The BRFSS dataset includes health-related survey data that is instrumental in understanding public health outcomes. We have collected datasets for 4 years.

```{r get-cdc-data}
library(haven)

cdc_data_2021 <- read_xpt("data/cdc/MMSA2021.xpt")
head(cdc_data_2021,5)
cdc_data_2020 <- read_xpt("data/cdc/MMSA2020.xpt")
head(cdc_data_2020,5)
cdc_data_2019 <- read_xpt("data/cdc/MMSA2019.xpt")
head(cdc_data_2019,5)
cdc_data_2018 <- read_xpt("data/cdc/MMSA2018.xpt")
head(cdc_data_2018,5)

```

```{r get-cdc-info}
# Listing columns for CDC datasets
cat("Columns for CDC 2021:\n", paste(names(cdc_data_2021), collapse = ", "), "\n\n")
cat("Columns for CDC 2020:\n", paste(names(cdc_data_2020), collapse = ", "), "\n\n")
cat("Columns for CDC 2019:\n", paste(names(cdc_data_2019), collapse = ", "), "\n\n")
cat("Columns for CDC 2018:\n", paste(names(cdc_data_2018), collapse = ", "), "\n\n")
```

## Feature Explanation

### AQS Dataset Features:

- **State Code, County Code, Site Num**: These identifiers allow us to link air quality data to specific geographic locations, enabling regional analysis and comparison.

- **Parameter Name**: This feature helps us focus on pollutants of interest, based on their known or suspected health impacts.

- **Arithmetic Mean, 1st Max Value**: Provide measures of average and peak pollutant levels, which are key for assessing exposure and potential health effects.

- **Year**: Allows for temporal analysis and matching with health data to observe trends over time.

### CDC Dataset Features:

- **General Health Indicators (GENHLTH, PHYSHLTH, MENTHLTH)**: Directly relate to the population's health status, serving as outcome variables in our analysis.

- **Demographic Information**: Enables the identification of populations potentially more vulnerable to poor air quality.

- **Chronic Conditions**: Offer insights into specific health outcomes that may be exacerbated by air pollution.

### Selection Criteria:

- We prioritize features that offer direct insights into air quality levels, health outcomes, and demographic characteristics relevant to our research questions.
- Features not selected lack a direct connection to our study objectives or are redundant with other data points.


## References



